---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load Data
```{r echo=TRUE}
data <- read.csv("data/activity.csv")
```

Preprocess Data - Remove all NAs
```{r echo=TRUE}
cleandata <- data[complete.cases(data),]
```

Preprocess Data - Change date datatype to date format
```{r echo=TRUE}
cleandata$date <- as.Date(cleandata$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Number of steps taken per day
```{r echo=TRUE}
totalSteps <- aggregate(steps ~ date, cleandata, sum)
```

Plot a histogram of the total number of steps taken each day
```{r echo=TRUE, message=FALSE}
library(ggplot2)
ggplot(totalSteps, aes(x=steps)) + geom_histogram() + xlab("Total Steps")
```

Calculate the mean of the total number of steps taken per day
```{r echo=TRUE}
meanSteps <- aggregate(steps ~ date, cleandata, mean)
meanSteps
```

Calculate the median of the total number of steps taken per day
```{r echo=TRUE}
medianSteps <- aggregate(steps ~ date, cleandata, median)
medianSteps
```

## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo=TRUE, message=FALSE}
dailyPattern <- aggregate(steps ~ interval, cleandata, mean)
ggplot(dailyPattern, aes(x=interval, y=steps)) +
        geom_line(color="blue", sie=1) +
        labs(title="Average Daily Acvity Pattern", x="Interval", y="Number of Steps") +
        theme_bw()
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
dailyPattern[dailyPattern$steps==max(dailyPattern$steps), ]
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r echo=TRUE}
incomplete_data <- !complete.cases(data)
nrow(data[incomplete_data,])
```

Devise a strategy for filling in all of the missing values in the dataset.
- To use average interval value to replace the NAs
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r echo=TRUE}
meandata <- data
nas <- is.na(meandata$steps)
avg_interval <- tapply(meandata$steps, meandata$interval, mean, na.rm=TRUE, simplify=TRUE)
meandata$steps[nas] <- avg_interval[as.character(meandata$interval[nas])]
```

Make a histogram of the total number of steps taken each day (after NAs value replaced)
```{r echo=TRUE, message=FALSE}
totalSteps2 <- aggregate(steps ~ date, meandata, sum)
ggplot(totalSteps2, aes(x=steps)) + geom_histogram() + xlab("Total Steps")
```

Calculate and report the mean total number of steps taken per day.
```{r echo=TRUE}
meanSteps2 <- aggregate(steps ~ date, meandata, mean)
meanSteps2
```

Calculate and report the median total number of steps taken per day.
```{r echo=TRUE}
medianSteps2 <- aggregate(steps ~ date, meandata, median)
medianSteps2
```

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
day <- weekdays(as.Date(cleandata$date))
daylevel <- vector()
for (i in 1:nrow(cleandata)) {
        if (day[i] == "Saturday") {
                daylevel[i] <- "Weekend"
        }        
        else if (day[i] == "Sunday") {
                daylevel[i] <- "Weekend"
        } 
        else {
                daylevel[i] <- "Weekday"
        }
}
cleandata$daylevel <- daylevel
cleandata$daylevel <- factor(cleandata$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = cleandata, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
library("lattice")
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```